Carbonate chemistry at Ny-Ålesund

### 2021-05-14 awipev-CO2.Rmd
finalized co2_air

```{r atmospheric CO2 Zeppelin, echo=FALSE, message= FALSE, warning=FALSE}
## Atmospheric CO2 Zeppelin
# downloaded 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999
# value:units : ppm
# value:long_name : measured_mole_fraction_of_trace_gas_in_dry_air
atm_co2_old <-
  read_delim(file = paste0(path, "atmospheric_co2_zeppelin_old/co2_zep_surface-insitu_54_9999-9999_hourly.txt"),
  delim = " ",
  comment = "#",
  #skip = 182,
  col_names = TRUE,
  na = "-999.999",
  ) %>%
  dplyr::mutate(co2_air = value,
  datetime = as_datetime(paste(paste(year, month, day, sep="-"), paste(hour, minute, second, sep="-"), sep = " "))) %>%
  dplyr::select(datetime, co2_air)

# downloaded 2021-04-14
# http://ebas.nilu.no/Pages/DataSetList.aspx?key=3673A72690B343B1AEC0F1673FE8C09D
file_names <- c("atmospheric_co2_zeppelin/NO0042G.20160101000000.20180420115103.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20170101000000.20180413180000.online_crds.carbon_dioxide.air.7mo.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", 
"atmospheric_co2_zeppelin/NO0042G.20180101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20190101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20200101000000.20210413000000.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas")
atm_co2_new <- NULL
year_co2 <- ymd_hm(paste0(c(2016:2020), "-01-01 00:00"))
for (i in 1:5) {
print(i)
tmp <-
  read_table2(file = paste0(path, file_names[i]),
  comment = "#",
  #skip = 182,
  col_names = FALSE,
  na = c("9999.999", "99.999", "9999.990", "9999.99"),
  ) %>%
  dplyr::mutate(co2_air = X3,
  datetime = year_co2[i] + X1*24*3600) %>% 
  dplyr::select(datetime, co2_air)
  tmp$datetime <- lubridate::round_date(tmp$datetime, unit="1 minutes")
atm_co2_new <- bind_rows(atm_co2_new, tmp)
  }
# combine the two sources of CO2 data
# the new source has missing data in 2018, so the data from the old source
# are used for 2018
atm_co2 <- dplyr::left_join(atm_co2_new, atm_co2_old, by = "datetime")
atm_co2 <- atm_co2 %>% 
  dplyr::rename(co2_air = co2_air.x, co2_air_old = co2_air.y) %>%
  dplyr::mutate(co2_air = ifelse(year(atm_co2$datetime) == 2018, co2_air_old, co2_air))
# summary(atm_co2)
zz <- as_tsibble(atm_co2) %>% 
  mutate(smoothed_co2_air = slide_dbl(co2_air, .f = mean,
                                   .before = 1, .after = 1))
ggplot(zz) +
  aes(x = time_hour, y = smoothed_temp, color = origin)

slide_dbl(~mean(.x$datetime, .x$co2_air), .before = 3, .after = 3)
  #slide_period_dbl(.x = atm_co2$co2_air, .f = mean, .i = atm_co2$datetime, .period = "month")

#d_all <- dplyr::left_join(d_all, atm_co2, by = "datetime")
p_atm_co2 <- atm_co2 %>% 
  #tidyr::pivot_longer(cols = c(pco2_air, pco2_air_old), names_to = "dataset", values_to = "pco2") %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = running_mean(atm_co2$co2_air, 24)), col = "blue", size = 0.15, na.rm = TRUE) +
  #facet_grid(rows = vars(dataset)) +
  #geom_point(aes(x = datetime, y = pco2_air_old), col = "grey", size = 0.15, na.rm = TRUE) +
  labs(title = "Atmospheric CO2 at Zeppelin", x = "Time", y = "CO2 (ppm)") +
  Mytheme(size_labs = 8)
p_atm_co2
ggsave(filename = "figures/atm_co2.png", plot = p_atm_co2, width = 20, units = "cm")
```

@@

load(file = paste0(path, "meteo_awi/wind_nyalesund.RData")) 
meteo <- tibble(windnyalesund) %>%
  dplyr::mutate(datetime = as_datetime(paste(paste(Year, Month, Day, sep="-"), paste(Hour, "00", "00", sep="-"), sep = " ")),
         temp_air = `Temperature[degC]`,
         p_air = `Pressure[hPa]`,
         ws = `WindSpeed[m/s]`,
         wd = `WindDirection[deg]`
        ) %>%
  dplyr::select(datetime, temp_air, p_air, ws, wd)
d_all <- dplyr::left_join(d_all, meteo, by = "datetime") %>% 
  select(sort(dplyr::everything()))
```

>>>>

```{r atmospheric CO2 Zeppelin, echo=FALSE, message= FALSE, warning=FALSE}
## Atmospheric CO2 Zeppelin
# downloaded 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999
# value:units : ppm
# value:long_name : measured_mole_fraction_of_trace_gas_in_dry_air
atm_co2_old <-
  read_delim(file = paste0(path, "atmospheric_co2_zeppelin/co2_zep_surface-insitu_54_9999-9999_hourly.txt"),
  delim = " ",
  comment = "#",
  #skip = 182,
  col_names = TRUE,
  na = "-999.999",
  ) %>%
  dplyr::mutate(co2_air = value,
  datetime = as_datetime(paste(paste(year, month, day, sep="-"), paste(hour, minute, second, sep="-"), sep = " "))) %>%
  dplyr::select(datetime, co2_air)

# downloaded 2021-04-14
# http://ebas.nilu.no/Pages/DataSetList.aspx?key=3673A72690B343B1AEC0F1673FE8C09D
file_names <- c("atmospheric_co2_zeppelin/NO0042G.20160101000000.20180420115103.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20170101000000.20180413180000.online_crds.carbon_dioxide.air.7mo.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", 
"atmospheric_co2_zeppelin/NO0042G.20180101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20190101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20200101000000.20210413000000.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas")
atm_co2_new <- NULL
year_co2 <- ymd_hm(paste0(c(2016:2020), "-01-01 00:00"))
for (i in 1:5) {
tmp <-
  read_table2(file = paste0(path, file_names[i]),
  comment = "#",
  #skip = 182,
  col_names = FALSE,
  na = c("9999.999", "99.999", "9999.990", "9999.99"),
  ) %>%
  dplyr::mutate(co2_air = X3,
  datetime = year_co2[i] + X1*24*3600) %>% 
  dplyr::select(datetime, co2_air)
  tmp$datetime <- lubridate::round_date(tmp$datetime, unit="1 minutes")
atm_co2_new <- bind_rows(atm_co2_new, tmp)
  }
# combine the two sources of CO2 data
# the new source has missing data in 2018, so the data from the old source
# are used for 2018
atm_co2 <- dplyr::full_join(atm_co2_new, atm_co2_old, by = "datetime")
atm_co2 <- atm_co2 %>% 
  dplyr::rename(co2_air_new = co2_air.x, co2_air_old = co2_air.y) %>%
  dplyr::mutate(co2_air = ifelse(year(atm_co2$datetime) == 2020, co2_air_new, co2_air_old))
# Below was an attempt to smooth but
# zz <- as_tsibble(atm_co2) %>% 
#   mutate(smoothed_co2_air = slide_dbl(co2_air, .f = mean, .before = 12, .after = 12, .complete = TRUE))
# ggplot(zz) +
#   geom_line(aes(x = datetime, y = smoothed_co2_air))
d_all <- dplyr::left_join(d_all, dplyr::select(atm_co2, datetime, co2_air), by = "datetime") %>% 
  select(sort(current_vars()))

p_atm_co2 <- d_all %>% 
  #tidyr::pivot_longer(cols = c(pco2_air, pco2_air_old), names_to = "dataset", values_to = "pco2") %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = co2_air), col = "blue", size = 0.15, na.rm = TRUE) +
  #facet_grid(rows = vars(dataset)) +
  #geom_point(aes(x = datetime, y = pco2_air_old), col = "grey", size = 0.15, na.rm = TRUE) +
  labs(title = "Atmospheric CO2 at Zeppelin", x = "Time", y = "CO2 (ppm)") +
  Mytheme(size_labs = 8)
#p_atm_co2
ggsave(filename = "figures/atm_co2.png", plot = p_atm_co2, width = 20, units = "cm")
```

@@

load(file = paste0(path, "meteo_awi/wind_nyalesund.RData")) 
meteo <- tibble(windnyalesund) %>%
  dplyr::mutate(datetime = as_datetime(paste(paste(Year, Month, Day, sep="-"), paste(Hour, "00", "00", sep="-"), sep = " ")),
         temp_air = `Temperature[degC]`,
         p_air = `Pressure[hPa]`,
         ws = `WindSpeed[m/s]`,
         wd = `WindDirection[deg]`
        ) %>%
  dplyr::select(datetime, temp_air, p_air, ws, wd)
zz <- dplyr::left_join(d_all, meteo, by = "datetime") %>% 
  select(sort(current_vars()))
```
